---
title: "p8105_hw2_tw2918_01"
output: github_document
date: "2023-09-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)
```

# Probelm 1
## 1. Clean the data in pols-month.csv.
### (1) Use `separate()` to break up the variable `mon` into integer variables `year`, `month`, and `day`

### (2)  replace month number with month abbreviation

### (3) create a `president` variable taking values `gop` and `dem`, and remove `prez_dem` and `prez_gop`

### (4) remove `day` variable 

```{r, eval=FALSE}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )
```

```{r, eval=FALSE}
pols_month_df = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE)|>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -month_abb, -day, -starts_with("prez"))
```
 
## 2. clean the data in snp.csv using a similar process to the above
### (1) Use `separate()` to break up the variable `date` into integer variables `year`, `month`, and `day`
### (2)  replace month number with month abbreviation and remove `day` varialbe
### (3)  arrange according to year and month

```{r, eval=FALSE}
snp_df = 
  read_csv("./fivethirtyeight_datasets/snp.csv")|>
  janitor::clean_names()|>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month)|>
  mutate(month = month.name[as.numeric(month)])|>
  select(year, month,close)
```

## 3. tidy the unemployment data

```{r, eval=FALSE}
unemp_df = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv")|>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

## 4. Join the datasets by merging `snp` into `pols`, and merging `unemployment` into the result

```{r}
```

# Probelm 2 
## 1. Read and Clearn Mr. Trash Wheel Dataset

```{r}
mr_trashwheel_df = 
  read_excel(
    "./202207 Trash Wheel Collection Data.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = "A2:N549"
    ) |>
  janitor::clean_names()

```


