---
title: "p8105_hw2_tw2918_01"
output: github_document
date: "2023-09-30"
---

```{r}
library(tidyverse)
```

# Probelm 1
## 1. Clean the data in pols-month.csv.
### (1) Use `separate()` to break up the variable `mon` into integer variables `year`, `month`, and `day`

### (2)  replace month number with month abbreviation

### (3) create a `president` variable taking values `gop` and `dem`, and remove `prez_dem` and `prez_gop`

### (4) remove `day` variable 

```{r, eval=FALSE}
pols_month_df = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into = c("year", "month", "day"))|>
  mutate(month = month.abb[as.numeric(month)])|>
  relocate(year, month, day, prez_gop, prez_dem) |>
  pivot_longer(
    prez_gop: prez_dem,
    names_to= "president",
    names_prefix = "prez_",
  )|>
  select(-value, -day)
```

## 2. clean the data in snp.csv using a similar process to the above
### (1) Use `separate()` to break up the variable `date` into integer variables `year`, `month`, and `day`
### (2)  replace month number with month abbreviation and remove `day` varialbe
### (3)  arrange according to year and month

```{r, eval=FALSE}
snp_df = 
  read_csv("./fivethirtyeight_datasets/snp.csv")|>
  janitor::clean_names()|>
  separate(date, into = c("month", "day", "year"))|>
  select(-day)|>
  mutate(month = month.abb[as.numeric(month)])|>
  arrange(year, month)|>
  relocate(year, month,close)
```